A Comparative study of Dask and Spark for data-intensive neuroimaging pipelines

### 2019-07-11 latex/paper.tex
Wroked on comments

 We used Compute Canada's
 \href{https://docs.computecanada.ca/wiki/Cloud_resources}{Arbutus Cloud}
 operated by the \href{https://www.westgrid.ca}{WestGrid} regional
 organization at the University of Victoria, and running OpenStack
 \TG{version?}. We used c8-30gb-186 cloud instances with 8 VCPUs, an Intel
 Xeon Gold 6130 processor, 30~GB of RAM at 2666~MHz, 20~GB of mounted
 storage, and a base image running CentOS 7.5.1804 with  Linux kernel
 version
 3.10.0\-862.11.6.el7.x86\_64. Instances are connected by a 10Gb/s Ethernet network.

 Cloud instances hosted a single Dask or Spark worker, configured to use 8 CPUs. We
 used Dask's default configuration that uses all the available memory on the
 instance. Its default heuristic is to: target a 60\% memory load, spill to disk at
 70\%, pause the worker at 80\%, and terminate the worker at 95\%. We configured
 Spark to use 1 executor per worker and 25~GB of memory per executor, to leave 5~GB
 for off-heap. We configured the Spark driver to use 25~GB of memory, and used the
 default configuration for the master. We used the default confirguration for worker
 memory management: at 60\% it spill data to disk, and 50\% of that amount is
 reserved for storage and is immuned from eviction.

 One cloud instance did not host any worker, and had a 2~TB disk volume
 shared with the other instances using the Network File System (NFS) v4.
 This instance was also used for the Spark driver and master, and for the
 Dask scheduler, and for job monitoring with the Spark and Dask user
 interfaces. For both Spark and Dask, spilled data is evicted to the NFS.

@@

We also used the dataset provided by the Consortium for Reliability and
Reproducibility
(\href{http://fcon_1000.projects.nitrc.org/indi/CoRR/html/}{CoRR}) as
available on
\href{http://datasets.datalad.org/?dir=/corr/RawDataBIDS}{DataLad}. The
entire dataset is 408.4~GB, containing anatomical, diffusion and functional
images of 1,397 subjects acquired in 29 sites. Among these, we only used
data from \TG{X} subjects, representing \TG{Y}GB overall (\TG{Z}GB by
subject on average).
\MD{X, Y, Z values TBD after testing the makespan with different number of subjects.}

@@

We used three neuroimaging applications to evaluate the engines in different
conditions. The first two ones, incrementation and histogram, are simple synthetic
benchmarks representing basic map-only and map-reduce applications. The third one is
a realistic application representative of popular BIDS
applications~\cite{gorgolewski2017bids}.

@@

 \TG{Add links to the Github repo for applications and any other relevant file.}

@@

We used the \href{https://github.com/BIDS-Apps/example}{BIDS app example}
application available on GitHub. This application operates in a map-reduce
way. The map phase, also called participant analysis, extracts the brain
tissues of a subject's 3D MRI using the popular FMRIB Software Library
(\href{https://fsl.fmrib.ox.ac.uk/fsl/fslwiki}{FSL}), and writes the
resulting image (\TG{X}MB on average) to the shared file system. The reduce
phase, also called group analysis, computes the volume of each brain and
returns the average volume, shuffling a total of \TG{X}GB image data.
\MD{Determine values when dataset is determined for the application.}

While Incrementation and Histogram were implemented directly in Python,
BIDS app example requires an external library and is distributed as a
Docker container image (bids/base\_fsl on DockerHub). To avoid remote file
transfers from DockerHub, we preloaded our OpenStack instances with the
Docker image, and started the Docker engine at startup. \TG{explain run.py modifications if relevant.}

\subsection{Experiments}

@@

\begin{table}[!t]
    \renewcommand{\arraystretch}{1.3}
    \caption{Parameters for the experiments}\label{tab:param}
    \centering
    \begin{tabular*}{\columnwidth}{llll}
    \hline
                        & Incrementation & Histogram             & BIDS Example          \\ \hline
    \# of worker        & 1, 2, 4, 8     & 1, 2, 4, 8            & 1, 2, 4, 8            \\
    \# of blocks        & 30, 125, 750   & 30, 125, 750          & 30, 125, 750          \\
    \# of iterations    & 1, 10, 100     & \multicolumn{1}{c}{-} & \multicolumn{1}{c}{-} \\
    Sleep delay {[}s{]} & 1, 4, 16, 64   & \multicolumn{1}{c}{-} & \multicolumn{1}{c}{-} \\ \hline
    \end{tabular*}
    \end{table}

@@

%%% INCREMENTATION %%%
%% worker
\subsection{Incrementation: Number of workers}
Figure~\ref{fig:inc_ms_worker} shows the makespan of the Incrementation
application for different numbers of workers and for different engines. 
The bars show the average makespan over 5 repetitions while the error bars 
are the standard deviation \TG{check this}.
Overall, there is no substantial makespan difference between the engines.
Dask seems to have a slight advantage over Spark, \TG{X (toutes les Dask vs Spark)} seconds on
average, with Delayed and Futures being slightly better than Bags.

@@

\begin{figure}[!b]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_worker.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_worker}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_worker.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_worker}
    \end{subfigure}
    \caption{125 blocks, 10 iterations, 4-s delay, 8 threads/worker}\label{fig:inc_worker}
\end{figure}

@@

Figure~\ref{fig:inc_ms_block} shows the Incrementation makespan when
varying the number of image blocks for a constant BigBrain image size. We
were not able to run Spark for 30 blocks due to its 2GB limitation in the
task size it can compute. Once again, we do not observe any significant
difference among the engines. For all engines, makespan variability
increases with the number of blocks, however, engines scale very well in
general.

@@

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_block.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_block}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_block.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_block}
    \end{subfigure}
    \caption{10 iterations, 4-s delay, 8 workers, 8 threads/worker}\label{fig:inc_block}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_itr.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_itr}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_itr.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_itr}
    \end{subfigure}
    \caption{125 blocks, 4-s delay, 8 workers, 8 threads/worker}
\end{figure}

@@

Figure~\ref{fig:inc_gantt} shows the Gantt chart obtained for each engine
and API, for a baseline execution with 125 blocks, a 4-second sleep delay
and 8 workers \TG{Make the colors uniform}. Gantt charts are structured in 
batches of up to 64 read-compute-write concurrent sequences.
File reads in the first batch are much longer than in the following ones:
this is due to the high synchronization of data transfers that leads to a
high saturation of the shared file system. We also note that overhead,
represented in white, is concentrated around the data transfer tasks
\TG{find out why}, and the computing tasks that run concurrently with data
transfers. 
% The frameworks have similar makespan however the time spend for each type of
% task differs considerably (see Table~\ref{tab:inc_base}). Spark tends to spend more
% time than other frameworks for IO however it has a much lower overhead. Dask Delayed
% and Futures have the lowest IO time but a much higher overhead. Dask Bag stands in
% the middle for both IO and overhead. Compute time is approximately the same for all
% frameworks. Apart from the first IO section, there is a large amount of ovehead
% everytime time IO is performed. This is true for all frameworks.

\begin{figure*}[!htb]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/spark_inc_baseline_gantt.png}
        \caption{Spark execution timeline}\label{fig:inc_spark_gantt}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/Dask_bag_inc_baseline_gantt.png}%
        \caption{Dask Bag execution timeline}\label{fig:inc_dask_bag_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/Dask_delayed_inc_baseline_gantt.png}%
        \caption{Dask Delayed execution timeline}\label{fig:inc_dask_delayed_gantt}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/Dask_futures_inc_baseline_gantt.png}%
        \caption{Dask Futures execution timeline}\label{fig:inc_dask_futures_gantt}
    \end{subfigure}
    \caption{125 blocks, 4 sec.\ sleep delay, 8 workers}\label{fig:inc_gantt}
\end{figure*}

@@

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_instance.png}%
        \caption{Histogram makespan}\label{fig:histo_ms_worker}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_idle_instances.png}%
        \caption{Histogram total time}\label{fig:histo_tt_worker}
    \end{subfigure}
    \caption{125 blocks}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_splits.png}%
        \caption{Histogram makespan}\label{fig:histo_ms_block}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_idle_splits.png}%
        \caption{Histogram total time}\label{fig:histo_tt_block}
    \end{subfigure}
    \caption{8 workers}
\end{figure}
\TG{We decided to omit Dask Futures as it does not have any value over Dask
Delayed in this application.}
%%% BIDS EXAMPLE %%%
\subsection{Experiment 3: Number of workers}

@@

\subsection{IO Time}

\subsection{Serialization}
We think that serialization could have an important effect on the makespan of the
application. This could potentially slow down the application when a lot of tasks are
scheduled.

\subsection{NFS}
The NFS seems to be a source of the bottleneck. We think that exploring another file
system, like Lustre, could give us insight into the actual effect of the NFS on the
application makespan. We also think that the NFS caching could play an effect on the
task of different lengths.

\subsection{Scheduler}
The Dask scheduler seems to have some bizarre behavior. It seems like it waits to
schedule tasks. It could also be due to tasks being scheduled on other workers
thus requiring data to be sent over the network; which would explain the stall on the
worker. More work would be needed to investigate that issue.

\subsection{Caching}
We think that caching plays an effect on the results. As seen in the Gantt chart
previously, some of the read tasks are significantly shorter than others while all
blocks are of equal size.

>>>>

 We used Compute Canada's
 \href{https://docs.computecanada.ca/wiki/Cloud_resources}{Arbutus Cloud} operated by
 the \href{https://www.westgrid.ca}{WestGrid} regional organization at the University
 of Victoria, and running OpenStack \TG{version?}. We used c8-30gb-186 cloud
 instances with 8 VCPUs, an Intel Xeon Gold 6130 processor, \SI{30}{\giga\byte} of
 RAM at \SI{2666}{\mega\hertz}, \SI{20}{\giga\byte} of mounted storage, and a base
 image running CentOS 7.5.1804 with Linux kernel version
 3.10.0\-862.11.6.el7.x86\_64. Instances are connected by a
 \SI{10}{\giga\bit/\second} Ethernet network.

 Cloud instances hosted a single Dask or Spark worker, configured to use 8 CPUs. We
 used Dask's default configuration that uses all the available memory on the
 instance. Its default heuristic is to: target a 60\% memory load, spill to disk at
 70\%, pause the worker at 80\%, and terminate the worker at 95\%. We configured
 Spark to use 1 executor per worker and \SI{25}{\giga\byte} of memory per executor, to leave 5~GB
 for off-heap. We configured the Spark driver to use 25~GB of memory, and used the
 default configuration for the master. We used the default confirguration for worker
 memory management: at 60\% it spill data to disk, and 50\% of that amount is
 reserved for storage and is immuned from eviction.

 One cloud instance did not host any worker, and had a \SI{2}{\tera\byte} disk volume
 shared with the other instances using the Network File System (NFS) v4.
 This instance was also used for the Spark driver and master, and for the
 Dask scheduler, and for job monitoring with the Spark and Dask user
 interfaces. For both Spark and Dask, spilled data is evicted to the NFS.

@@

We also used the dataset provided by the Consortium for Reliability and
Reproducibility (\href{http://fcon_1000.projects.nitrc.org/indi/CoRR/html/}{CoRR}) as
available on \href{http://datasets.datalad.org/?dir=/corr/RawDataBIDS}{DataLad}. The
entire dataset is \SI{408.4}{\giga\byte}, containing anatomical, diffusion and
functional images of 1,397 subjects acquired in 29 sites. We used all 3491 anatomical
images, representing \SI{39}{\giga\byte} overall (\SI{11.17}{\mega\byte} by image on
average).

@@

We used three neuroimaging applications to evaluate the engines in different
conditions. The first two ones, incrementation and histogram, are simple synthetic
benchmarks representing basic map-only and map-reduce applications. The third one is
a realistic application representative of popular BIDS
applications~\cite{gorgolewski2017bids}. All script used for our experiment is availble
GitHub at \url{https://github.com/big-data-lab-team/paper-big-data-engines}

@@

@@

We used the \href{https://github.com/BIDS-Apps/example}{BIDS app example} application
available on GitHub. This application operates in a map-reduce way. The map phase,
also called participant analysis, extracts the brain tissues of a subject's 3D MRI
using the popular FMRIB Software Library
(\href{https://fsl.fmrib.ox.ac.uk/fsl/fslwiki}{FSL}), and writes the resulting image
(\SI{2.47}{\mega\byte} on average) to the shared file system. The reduce phase, also
called group analysis, computes the volume of each brain and returns the average
volume, shuffling a total of \SI{8.6}{\giga\byte} image data.

While Incrementation and Histogram were implemented directly in Python,
BIDS app example requires an external library and is distributed as a
Docker container image (bids/base\_fsl on DockerHub). To avoid remote file
transfers from DockerHub, we preloaded our OpenStack instances with the
Docker image, and started the Docker engine at startup.
\subsection{Experiments}

@@

\begin{table}[!t]
    \renewcommand{\arraystretch}{1.3}
    \caption{Parameters for the experiments}\label{tab:param}
    \centering
    \begin{tabular*}{\columnwidth}{llll}
    \hline
                        & Incrementation & Histogram             & BIDS Example          \\ \hline
    \# of worker        & 1, 2, 4, 8     & 1, 2, 4, 8            & 1, 2, 4, 8            \\
    \# of blocks        & 30, 125, 750   & 30, 125, 750          & 30, 125, 750          \\
    \# of iterations    & 1, 10, 100     & \multicolumn{1}{c}{-} & \multicolumn{1}{c}{-} \\
    Sleep delay {[}\SI{}{\second}{]} & 1, 4, 16, 64   & \multicolumn{1}{c}{-} & \multicolumn{1}{c}{-} \\ \hline
    \end{tabular*}
    \end{table}

@@

%%% INCREMENTATION %%%
%% worker
\subsection{Incrementation: Number of workers}
Figure~\ref{fig:inc_ms_worker} shows the makespan of the Incrementation application
for different numbers of workers and for different engines. The bars show the average
makespan over 5 repetitions while the error bars are the standard deviation. Overall,
there is no substantial makespan difference between the engines. Dask seems to have a
slight advantage over Spark, \SI{83.61}{\second} on average,
with Delayed and Futures being slightly better than Bags.

@@

\begin{figure}[!b]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_worker.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_worker}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_worker.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_worker}
    \end{subfigure}
    \caption{125 blocks, 10 iterations, \SI{4}{\second} delay, 8 threads/worker}\label{fig:inc_worker}
\end{figure}

@@

Figure~\ref{fig:inc_ms_block} shows the Incrementation makespan when varying the
number of image blocks for a constant BigBrain image size. We were not able to run
Spark for 30 blocks due to its \SI{2}{\giga\byte} limitation in the task size it can
compute. Once again, we do not observe any significant difference among the engines.
For all engines, makespan variability increases with the number of blocks, however,
engines scale very well in general.

@@

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_block.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_block}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_block.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_block}
    \end{subfigure}
    \caption{10 iterations, \SI{4}{\second} delay, 8 workers, 8 threads/worker}\label{fig:inc_block}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_itr.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_itr}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_itr.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_itr}
    \end{subfigure}
    \caption{125 blocks, \SI{4}{\second} delay, 8 workers, 8 threads/worker}
\end{figure}

@@

Figure~\ref{fig:inc_gantt} shows the Gantt chart obtained for each engine and API,
for a baseline execution with 125 blocks, a \SI{4}{\second} sleep delay and 8
workers. Gantt charts are structured in batches of up to 64 read-compute-write
concurrent sequences. File reads in the first batch are much longer than in the
following ones: this is due to the high synchronization of data transfers that leads
to a high saturation of the shared file system. We also note that overhead,
represented in white, is concentrated around the data transfer tasks \TG{find out
why}, and the computing tasks that run concurrently with data transfers. 
% The frameworks have similar makespan however the time spend for each type of
% task differs considerably (see Table~\ref{tab:inc_base}). Spark tends to spend more
% time than other frameworks for IO however it has a much lower overhead. Dask Delayed
% and Futures have the lowest IO time but a much higher overhead. Dask Bag stands in
% the middle for both IO and overhead. Compute time is approximately the same for all
% frameworks. Apart from the first IO section, there is a large amount of ovehead
% everytime time IO is performed. This is true for all frameworks.

\begin{figure*}[!htb]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/spark_inc_gantt.png}
        \caption{Spark execution timeline}\label{fig:inc_spark_gantt}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/bag_inc_gantt.png}%
        \caption{Dask Bag execution timeline}\label{fig:inc_dask_bag_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/delayed_inc_gantt.png}%
        \caption{Dask Delayed execution timeline}\label{fig:inc_dask_delayed_gantt}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/futures_inc_gantt.png}%
        \caption{Dask Futures execution timeline}\label{fig:inc_dask_futures_gantt}
    \end{subfigure}
    \caption{125 blocks, 4 sec.\ sleep delay, 8 workers}\label{fig:inc_gantt}
\end{figure*}

@@

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_worker.png}%
        \caption{Histogram makespan}\label{fig:histo_ms_worker}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_idle_worker.png}%
        \caption{Histogram total time}\label{fig:histo_tt_worker}
    \end{subfigure}
    \caption{125 blocks}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_block.png}%
        \caption{Histogram makespan}\label{fig:histo_ms_block}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_idle_block.png}%
        \caption{Histogram total time}\label{fig:histo_tt_block}
    \end{subfigure}
    \caption{8 workers}
\end{figure}
\TG{We decided to omit Dask Futures as it does not have any value over Dask
Delayed in this application.}
%%% BIDS EXAMPLE %%%
\subsection{Experiment 3: Number of workers}

@@

% \subsection{IO Time}
% 
% 
% \subsection{Serialization}
% We think that serialization could have an important effect on the makespan of the
% application. This could potentially slow down the application when a lot of tasks are
% scheduled.
% 
% \subsection{NFS}
% The NFS seems to be a source of the bottleneck. We think that exploring another file
% system, like Lustre, could give us insight into the actual effect of the NFS on the
% application makespan. We also think that the NFS caching could play an effect on the
% task of different lengths.
% 
% \subsection{Scheduler}
% The Dask scheduler seems to have some bizarre behavior. It seems like it waits to
% schedule tasks. It could also be due to tasks being scheduled on other workers
% thus requiring data to be sent over the network; which would explain the stall on the
% worker. More work would be needed to investigate that issue.
% 
% 
% \subsection{Caching}
% We think that caching plays an effect on the results. As seen in the Gantt chart
% previously, some of the read tasks are significantly shorter than others while all
% blocks are of equal size.
